Universals of word order result from optimization of grammars for efficient communication

### 2019-03-15 second-draft/3-paper-names.tex
revising

\dropcap{F}or TODO rewrite first sentence. decades, researchers in fields ranging from philology to cognitive science to statistical physics have been involved in documenting and trying to explain the universal syntactic and statistical properties of human language \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical}. An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}. In this work we examine this question from a computational perspective, demonstrating a fully formalized framework in which certain syntactic universals can be explained through the statistical optimization of grammars for information-theoretic efficiency. % TODO add a cogsci person to citation chain at the beginning

@@

Natural languages vary a lot in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  `letter', and a phrase translating to `to a friend'.
In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
While the two sentences containing words with the same meanings, the order of these words are entirely different in the two languages:
%However, the languages differ in the order in which these words and phrases are arranged: 
In Arabic, the verb stands at the beginning, followed by both the object and the adpositional phrase.
`To' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
In Japanese, the verb stands at the end; object and adpositional phrase precede it; `to' is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
%It turns out that these are deep and stable regularities:
It turns out that this variation reflects a deep and stable regularity:
While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
this is strongly correlated with the occurrence of pre- or postpositions (Figure~\ref{fig:corr-table}, top):
Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
%However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
%As Figure~\ref{fig:corr-table} shows, this correlation has an extremely strong empirical basis:
%In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization. %, over 90 \% of languages satisfy this generalization .

@@

Consequently, explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}.

@@

Prior studies of word order typology have proposed Dependency Length Minimization (DLM) as general principle of word order \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
This is the idea that languages order words so as to minimize the linear distance between words linked in grammatical relationships.
This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in language processing \cite{gibson1998linguistic,gibson2000dependency}.
Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
It has also been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.
As a principle of word order, DLM is also stipulative: There are other processing models, such as surprisal, that are strongly predictive of human processing cost, and make predictions distinct from, and partly conflicting with, those of DLM~\cite{hale2001probabilistic,levy2008expectation}.

@@

When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
The listener receives this string of words.
By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).

@@

In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes.
In our model, these will depend on the grammar:
Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
This reflects the \key{parseability} of the language.

@@

We formalize the cost of an utterance as its word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
Surprisal has been found to be is a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
%The \key{predictability} of the language is the negative entropy of the sentences:
\begin{equation}
	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
\end{equation}
where the sum runs over all possible sentences $l$ that belong to the language.
In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. %how many distinct sentences there are.

@@

Our goals are to show that
(1) the grammars of natural languages evolve towards optimizing efficiency of communication, and
(2) this process of optimization accounts both for DLM and for Greenberg's order correlation universals.

@@

We first demonstrate that real languages are relatively efficient compared to random baselines. For each language, we generated ten baseline word order grammars for the language.
These grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}).
The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.

@@

%\begin{figure*}
%\includegraphics[width=0.98\textwidth]{../results/plot-by-lang/grid-optimized.pdf} % need MLE langs and Efficiency langs on here
%	\caption{Predictability (top) and parseability (bottom) of real, optimized, and random languages, for English (left) and Japanese (right). We provide the entropy $H[L_\theta(T)]$ for predictability (lower is better) and the conditional entropy $H[T|L_\theta(T)]$ for parseability (lower is better). Both measures are normalized by sentence length.}\label{fig:predict-pars}
%\end{figure*}
%

\subsection{Efficiency Explains DLM}
We demonstrate the relationship between dependency length minimization and the maximization of efficiency. Figure~\ref{fig:deplength} shows average dependency length per sentence length for four typologically distinct languages, showing real languages, random baselines, and languages optimized for dependency length, parseability, predictability, and efficiency. We see that optimizing for efficiency lowers dependency length relative to random baselines, confirming the suggestion that dependency length minimization is a by-product of efficiency maximization \cite{futrell2017memory}. In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; on average, the real language is best matched by efficiency optimization.

@@

	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{DepL}   &  \multirow{2}{*}{Eff}  \\ 

@@

	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_case.pdf}       }

@@

	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_cop.pdf}       }

@@

	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-aux.pdf}       }

@@

	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-nmod.pdf}     }

@@

	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-acl.pdf}     }

@@

	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-lifted_mark.pdf}     }

@@

	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/bars-depl-obl.pdf}     }

@@

	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.

@@

	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  &  \multirow{2}{*}{Eff}  \\ 
	& \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}       }
%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}   }   
	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}     }  \\
	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}       }
%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}   }   
	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}   }  \\
	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}       }
%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}     }
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  }  \\
	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  }  \\
	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}     }
%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf}  }   
	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  }  \\
	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf}     }
%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}     }
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf} }   
	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  }  \\
	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf}     }
%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}     }
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  }  \\
	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}     }
%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
	&   \parbox[c]{0.05\textwidth}{\multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}  }}  \\
	\caption{Greenbergian Correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.

@@

These are shown in Table~\ref{table:corr-dryer}.
%Each correlation is visualized 
%
%
We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages. To evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks. %The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS. %in all except four languages, for WALS did not have data.

Table~\ref{table:corr-dryer} shows that grammars optimized for efficiency show correlations 
In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
We are interested in the direction and significance of this effect: 
If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.

\paragraph{Results}
Results are shown in Table~\ref{table:corr-dryer}.
All eight correlations are confirmed in the models estimated from the real orderings.
Efficiency significantly predicts all the word order correlations, each in the same direction as attested in the dependency corpora.
As efficiency predicts DLM, we predict that DLM also accounts for all dependencies.

We confirmed this finding by estimating the posterior distribution of the number of correlations that the optimization process correctly predicts.
The estimated posterior probability that efficiency optimization predicts less than all eight correlations is $0.0242$. The probability that it predicts less than seven of the correlations is $<10^{-4}$.
Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.

>>>>

\dropcap{F} \cite{behaghel1909beziehungen,zipf1949human,greenberg1963universals,lin2017critical} \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}

@@

Natural languages vary in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  `letter', and a phrase translating to `to a friend'.
In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
Yet, the order of these words are entirely different in the two languages:
%However, the languages differ in the order in which these words and phrases are arranged: 
In Arabic, the verb stands at the beginning, followed by the object and the adpositional phrase.
`To' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
In Japanese, the verb stands at the end; object and adpositional phrase precede it; `to' is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
%It turns out that these are deep and stable regularities:
This variation reflects a deep and stable regularity:
While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
this is strongly correlated with the occurrence of pre- or postpositions (Figure~\ref{fig:corr-table}, top):
Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
%However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
%As Figure~\ref{fig:corr-table} shows, this correlation has an extremely strong empirical basis:
%In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization. %, over 90 \% of languages satisfy this generalization .

@@

Explaining these patterns has been an important aim of linguistic research since Greenberg's seminal study.
An explanation for the universal properties of language would enable a deeper scientific understanding of what human language is and how to model it, with applications in psychology and natural language processing \cite{hawkins2007processing,bender2009linguistically,bender2013linguistic}. (TODO cite linguists)

@@

@@

When uttering, the speaker needs to choose an ordering in which to order the words in this tree to generate a string of words.
Should the speaker do it as
Mary saw John (like in English)
or as 
Mary John saw (like in Japanese)

Depending on the language, the listener decodes this.

The listener receives this string of words.
By the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined.
The dependency structure (including the labels on the arcs) specifies how the meanings of words are combined.
Therefore, a listener needs to recover the information provided in the dependency structure in order to understand a sentence correctly.
Consistent with Shannon's model, we assume a Bayesian listener who decodes dependency structures probabilistically~(Figure~\ref{fig:comm}).
If communication is successful, the listener can identify the intended structure with high confidence~(Figure~\ref{fig:comm}).

@@

In Information Theory, the usefulness of a communication channel depends on (1) how costly encoding and transmission is, and (2) how precisely messages can be recovered from codes.

In our model, these will depend on the grammar:
Grammars can differ in the degree to which a listener can unambiguously recover the dependency structure from an utterance.
This reflects the \key{parseability} of the language.

@@

We formalize the cost of an utterance as its word-by-word surprisal, the degree to which it is unpredictable from the general statistics of the language.
Surprisal has been found to be a highly accurate and general predictor human online processing difficulty \cite{hale2001probabilistic,levy2008expectation,smith2013effect}, and can be justified  in terms of predictive coding theories of information processing in the brain \cite{friston2009predictive} as well as the general algorithmic complexity of language generation and comprehension \cite{li2008introduction}.
In expectation over all utterances $u$ in a language, the negative surprisal describes the \key{predictability}, or negative entropy, of the utteranes: %this quantity is the entropy: % of the sentences:
%The \key{predictability} of the language is the negative entropy of the sentences:
\begin{equation}
	R_{Pred} := - H[\mathcal{U}] = \sum_{u} p(u) \log p(u)
\end{equation}
where the sum runs over all possible sentences $l$ that belong to the language.
In keeping with Zipf's Force of Unification, this quantity describes how homogeneous the language is, i.e., it is larger if the distribution over sentences is concentrated on a smaller number of frequent sentences. %how many distinct sentences there are.

@@

We examined whether
(1) the grammars of natural languages evolve towards optimizing efficiency of communication, and
(2) this process of optimization accounts both for Greenberg's order correlation universals.

@@

We first demonstrate that real languages are relatively efficient compared to random baselines. For each language, we generated ten baseline word order grammars for the language.
These grammars have systematic word order regularities similar to natural language, but do not exhibit any correlations among the orderings of different syntactic relations (compare Figure~\ref{fig:grammars}).
The majority of real languages in Figure~\ref{fig:pareto-plane} are below and to the left of their baseline equivalents, demonstrating that they are relatively high in predictability and/or parseability.
X languages improve over the baselines in parsability, Y improve in predictability.
Z improve in overall efficiency.
Pareto frontier

@@

@@

	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   &  \multirow{2}{*}{Eff}  \\ 

@@

@@

@@

@@

@@

@@

@@

@@

\caption{Efficiency optimization accurately predicts the Greenbergian correlations. Each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.

@@

These are shown in Table~\ref{table:corr-dryer}.
We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
To evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS. %in all except four languages, for WALS did not have data.

Table~\ref{table:corr-dryer} shows that grammars optimized for efficiency accurately predict correlations to hold at prevalences similar to actual languages.

%In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
We conducted a multivariate mixed-effects logistic regression model predicting which correlations an optimized grammar satisfies, entering the language and language family of the underlying corpus as random effects.\footnote{We coded language families according to \url{universaldependencies.org}.}
We are interested in the direction and significance of this effect: 
If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.

We estimated the posterior distribution of the number of correlations that the optimization process correctly predicts.
The estimated posterior probability that efficiency optimization predicts less than all eight correlations is $0.0242$. The probability that it predicts less than seven of the correlations is $<10^{-4}$.
%Predictability and parseability predict four and six correlations, respectively, making largely complementary predictions.
%\begin{tabular}{|ll|c|cc|}
%	\multicolumn{2}{|c|}{Correlates with...}   &         \multirow{2}{*}{Real}   & \multirow{2}{*}{DepL}  &  \multirow{2}{*}{Eff}  \\ 
%	& \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} } %\textbf{86}  }
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}       }
%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf} } %  \textbf{47 \%}$^{}$   }
%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}   }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf}     }  \\
%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} }% \textbf{94} }
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}       }
%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf} } % \textbf{53}     }
%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}   }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}   }  \\
%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}       }
%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}     }
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}  }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}  }  \\
%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}  }  \\
%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}     }
%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf}  }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}  }  \\
%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf}     }
%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}     }
%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf} }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}  }  \\
%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf}     }
%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}     }
%%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf} }   
%	&   \multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}  }  \\
%	&  \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf}     }
%%	&    \multirow{2}{*}{\includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}     }
%%	&   \multirow{2}{*}{ \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf}  }   
%	&   \parbox[c]{0.05\textwidth}{\multirow{2}{*}{  \includegraphics[width=0.05\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}  }}  \\

@@

\section{Alternative Theories}

Prior studies have proposed Dependency Length Minimization (DLM) as general principle of word order \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
This is the idea that languages order words so as to minimize the linear distance between words linked in grammatical relationships.
This principle is a heuristic formalization of the idea that long dependencies should create high memory requirements in language processing \cite{gibson1998linguistic,gibson2000dependency}.
Cross-linguistic corpus studies have shown that languages do minimize dependency length beyond chance, suggesting that DLM is a language universal \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
It has also been argued theoretically that this principle should account for Greenberg's correlations, though this link is nontrivial and depends on the precise distributions of tree topologies in syntactic structures.
As a principle of word order, DLM is also stipulative: There are other processing models, such as surprisal, that are strongly predictive of human processing cost, and make predictions distinct from, and partly conflicting with, those of DLM~\cite{hale2001probabilistic,levy2008expectation}.
\cite{futrell2017memory} suggests that efficiency predicts DLM.

We show that Efficiency optimization predicts DLM, providing a first-principles functional explanation.
Figure~\ref{fig:deplength} shows average dependency length for real languages, random baselines, and languages optimized for dependency length and efficiency.
Optimizing for efficiency lowers dependency length relative to random baselines.

Our results suggest that dependency length minimization is a by-product of efficiency maximization. In 80\% of the languages, optimizing explicitly for dependency length produces dependencies that overshoot the dependency length of the real language; on average, the real language is best matched by efficiency optimization.
